\LARGE \bfseries \vspace{-2cm} Burden and health-related quality of life among caregivers of people with motor neuron disease

### 2024-06-21 output/manuscript.tex
Updates from Overleaf

\usepackage{latex/ediArticle}  %from ediArticle.sty
%\usepackage{multirow}
\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage[normalem]{ulem}
%\usepackage[table]{xcolor}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Motor neuron disease (MND) is a progressive neurodegenerative disease that impacts not only patients but also their informal caregivers. Several studies on ALS in Australia \parencite{lillo_caregiver_2012}, United States \parencite{qutub_life_2014, burke_caregiver_2015, roach_dynamics_2009}, Turkey \parencite{tulek_care_2023}, Ireland \parencite{galvin_caregiving_2016}, Germany \parencite{schischlevskij_informal_2021}, and China \parencite{geng_patients_2017}, have investigated the factors contributing to caregiver burden in ALS and the associated impact of caregiving on the quality of life of these individuals. \textcite{lillo_caregiver_2012} found that patients' abnormal behavior and caregiver stress were the strongest predictors of high caregiver burden, while physical disability was not significantly associated.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}

\subsection{Study design and participants}
COMMEND was a XXX designed to YYY. The study design, patient
characteristics, and baseline costs and resource-use data
have been reported \parencite{gould_randomised_2022}. The present study focuses on the caregivers; the baseline characteristics and quality of life have been reported previously for patient  \parencite{gould_acceptance_2024}. Patients with
probable/confirmed?? MND, CRITERIA???, were enrolled between
DATE and DATE, mostly at WHERE. Patients were
also required??? to have a primary caregiver who was
willing to participate in the study. Caregivers were included if BRIEF INCLUSION CRITERIA [cite trial].

\subsection{Data and questionnaires}
Data were collected for patients and caregivers at the
baseline visit and at baseline, 6 months, and 18 months during WHICH?? visits. Full details of the baseline patient and caregiver
demographics and characteristics, including comorbidities
and medications used, have been reported previously \parencite{gould_acceptance_2024, keetharuth_costeffectiveness_2024}.

@@

\subsection{Statistical analysis}
We described the characteristics of the participants (patients with MND and caregivers) using means and standard deviations for continuous variables and using numbers and percentages for categorical variables. We used univariate linear regression to examine the direction and size of the relationships between caregiver burden and possible influencing factors on  caregivers’ burden (ZBI) and health-related quality of life (EQ-5D). We then conducted a multiple regression analysis to test the extent to which these variables explained ZBI and EQ-5D. We selected variables to add to multiple linear regression by adopting a critical level of significance (p $\leq$ 0.10) in the univariate analysis.
%Spearman correlation coefficients examined the within- subject associations between continuous variables such as caregiver ZBI total scores, EQ-5D index scores, EQ-5D VAS scores, and T-IADL, at baseline, at 18 months, and for the change from baseline to 18 months. We inter- preted correlation coefficients of 0–0.19 as very weak and 0.20–0.39 as weak.
All analysis were conducted using Stata version 18.0 (StataCorp, College Station, Texas, USA).

\section{Results}
At baseline, the study cohort analyzed comprised 93 caregivers that consent was obtained to participate in the study, alongside the patients in their care. Both ZBI and HRQoL data at was collected from a total of 86 caregivers at baseline, of which 85 had complete and data on all outcome measures. One caregiver-patient pair was were excluded from further analyses because the EQ-5D score was negative while the VAS score was 100, resulting in 84 full observations. %This represents about 1.5\% of an estimated 5700 patients (prevalence, 8.6 per 100,000 persons) that lived with MND in 2020 \textbf{10.1080/21678421.2019.1587629}. The total number of caregivers with available data at the 9-month timepoint (and for the change from baseline to 9 months was) was: ZBI, n = 66; EQ-5D, n = 67.

\subsection{Caregiver and patient characteristics}
\autoref{tab_desc_stat} summarizes the baseline sociodemographic and clinical characteristics of the caregivers and patients. Caregivers had a mean (SD) age of 59.7 (11.8) years and most (70.2\%) caregivers were female, consistent with previous literature \parencite{schischlevskij_informal_2021}. Approximately half of the caregivers (4.8\%) reported being in paid employment.  Most caregivers (92.9\%) were direct family members of the patients and likely living together with a partner with MND as observed in previous studies \parencite{burke_longitudinal_2018, galvin_caregiving_2016, schischlevskij_informal_2021}. The median duration of caregiving (DOC) was three hours per day (as stated by the main CG; additional hours by more than one CG per patient are possible).  The mean ZBI scores at baseline was 21.1 (12.3). The EQ-VAS mean (79.8/100) was the same as the estimated EQ-5D-5L index score with 0.80/1. These HRQoL values were \textbf{lower/higher} than in the general British population (EQ-VAS mean = XX.X/100, SD = X.X [REF], EQ-5D-5L index value mean = X.X, SD = X.X) [REF].

@@

\subsection{Impact of patient's functional status on caregivers}
The mean ZBI score by King's stage showed that caregivers patients with worse functional status had a greater burden, but results were not consistent. \autoref{fig:outcome-kings-stage} shows the ZBI score by King's stage at baseline, 6 months, and 9 months. The ZBI score was highest for carers of stage 4 patients, but higher for stage 2 compared to stage 3. Using baseline data only, scores were < 20 for stage 1/12 and >20 for stage 3/4. We found a weak, negative, correlation between the ZBI score and the ALSFRS-R ($\rho$ = −0.208, p < 0.002). Due to this inconsistency, functional status.

@@

\subsection{Caregiver HRQoL and burden over time}
The mean (SD) change in carers ZBI scores from baseline to 18 months was X.X (X.X) for the caregiver EQ-5D index score,X.X (X.X) for the caregiver EQ-5D VAS score, and X.X (X.X) for the ZBI score. For the two HRQoL measures, the change from baseline to 6 months or 9 months was not significant. We found significant differences in ZBI scores between 0 and 9  months (p value = X.XXX), mostly driven by the change ins cores beuween 0 and 6 months.

\autoref{fig:eq5d-scores-time} shows the EQ-5D domain scores over time. At baseline, no caregiver had any problem in self-care domain. %Examination of the EQ-5D domain scores (Fig. 1) showed that very few caregivers had extreme problems (<5\% for any domain) and the levels of problems were similar at baseline and at 18 months. At both time points, more caregivers had some problems in the domains of pain/discomfort (baseline: 42.8\%; 18 months: 46.8\%) and anxiety/depression (baseline: 30.3\%; 18 months: 33.9\%) than in the other three domains.

\begin{table}[H]
    \centering \singlespacing \small
    \caption{Means differences (95\% confidence intervals) and p-values from t-tests comparing of caregiver ZBI, EQ-VAS, and EQ-5D scores changes over time}
    \begin{tabular}{|L{6cm}|C{5cm}|C{2.5cm}|}
        \hline
        \PlainInput{tables/outcome-scores-time}
    \end{tabular}
    \label{outcome-scores-time}
    \caption*{\footnotesize 
                \textit{Notes:} CI, confidence interval; EQ-5D, EuroQol 5-dimension questionnaire; VAS, visual analog scale; ZBI, Zarit Burden Interview}
\end{table}

\hspace{1em}
\begin{figure}[H]
    \centering
    \includegraphics[width=1\linewidth]{figures/eq5d-domains-time.png}
    \caption{Problems in the caregiver EQ-5D domains at baseline, 6 months, and 9 months.}
    \label{fig:eq5d-scores-time}
    \caption*{\footnotesize \textit{Notes:} ANOVA, analysis of variance; EQ-5D, EuroQol 5-dimension questionnaire; VAS, visual analog scale; ZBI, Zarit Burden Interview}
\end{figure}

\subsection{Relationship between ZBI and EQ-5D}
Correlations were weak or very weak between EQ-5D (index and VAS scores) and ZBI at each time point and for the change scores (Table 3). For the ZBI, the 18-month change score correlation was −0.16 (95\% confidence interval, CI: −0.22, −0.09) for the EQ-5D VAS score and −0.09 (95\% CI: −0.15, −0.03) for the EQ- 5D index score.

\subsubsection{Factors influencing caregiver burden}
%As already shown above, the patient’s functional status had a clear influence on her/his caregiver’s burden. To provide a deeper analysis of further influencing factors on caregiver burden (ZBI score), multiple regression analysis was used (Table 3). The patient’s wheelchair dependency increased the ZBI score by 9.30 points. Additionally, a rise of 5.01 points was observed, if the patient needed permanent supervision. The main influence on the ZBI appeared to be the CG’s mental health impairment due to caregiving with an increase of 11.36 points. However, physical health impairment had no statistically significant impact on the ZBI score. Furthermore, patients’ age seemed to slightly lower the burden by −0.24 points for each increasing year. Neither the patient’s nor the CG’s gender had statistically significant impacts on the ZBI in the multivariate regression analysis.

\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% %%%
%%% Bibliography %%%

>>>>

\usepackage{latex/ediArticle}  %from ediArticle.sty
%\usepackage{multirow}
\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage[normalem]{ulem}
%\usepackage[table]{xcolor}
\usepackage{lineno}
\linenumbers

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Introduction}
Motor neuron disease (MND) is a progressive neurodegenerative disease that impacts not only patients but also their informal caregivers. Several studies on ALS in Australia \parencite{lillo_caregiver_2012}, United States \parencite{qutub_life_2014, burke_caregiver_2015, roach_dynamics_2009}, Turkey \parencite{tulek_care_2023}, Ireland \parencite{galvin_caregiving_2016}, Germany \parencite{schischlevskij_informal_2021}, and China \parencite{geng_patients_2017}, have investigated the factors contributing to caregiver burden in ALS and the associated impact of caregiving on the quality of life of these individuals. \textcite{lillo_caregiver_2012} found that patients' abnormal behavior and caregiver stress were the strongest predictors of high caregiver burden, while physical disability was not significantly associated.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Methods}

\subsection*{Study design and participants}
COMMEND was a XXX designed to YYY. The study design, patient
characteristics, and baseline costs and resource-use data
have been reported \parencite{gould_randomised_2022}. The present study focuses on the caregivers; the baseline characteristics and quality of life have been reported previously for patient  \parencite{gould_acceptance_2024}. Patients with
probable/confirmed?? MND, CRITERIA???, were enrolled between
DATE and DATE, mostly at WHERE. Patients were
also required??? to have a primary caregiver who was
willing to participate in the study. Caregivers were included if BRIEF INCLUSION CRITERIA [cite trial].

\subsection*{Data and questionnaires}
Data were collected for patients and caregivers at the
baseline visit and at baseline, 6 months, and 18 months during WHICH?? visits. Full details of the baseline patient and caregiver
demographics and characteristics, including comorbidities
and medications used, have been reported previously \parencite{gould_acceptance_2024, keetharuth_costeffectiveness_2024}.

@@

\subsection*{Statistical analysis}
We described the characteristics of the participants (patients with MND and caregivers) using means and standard deviations for continuous variables and using numbers and percentages for categorical variables. We used univariate linear regression to examine the direction and size of the relationships between caregiver burden and possible influencing factors on  caregivers’ burden (ZBI) and health-related quality of life (EQ-5D). We then conducted a multiple regression analysis to test the extent to which these variables explained ZBI and EQ-5D. We selected variables to add to multiple linear regression by adopting a critical level of significance (p $\leq$ 0.10) in the univariate analysis.
%Spearman correlation coefficients examined the within- subject associations between continuous variables such as caregiver ZBI total scores, EQ-5D index scores, EQ-5D VAS scores, and T-IADL, at baseline, at 18 months, and for the change from baseline to 18 months. We inter- preted correlation coefficients of 0–0.19 as very weak and 0.20–0.39 as weak.
All analysis were conducted using Stata version 18.0 (StataCorp, College Station, Texas, USA).

\section*{Results}
At baseline, the study cohort analyzed comprised 93 caregivers that consent was obtained to participate in the study, alongside the patients in their care. Both ZBI and HRQoL data at was collected from a total of 86 caregivers at baseline, of which 85 had complete and data on all outcome measures. One caregiver-patient pair was were excluded from further analyses because the EQ-5D score was negative while the VAS score was 100, resulting in 84 full observations. %This represents about 1.5\% of an estimated 5700 patients (prevalence, 8.6 per 100,000 persons) that lived with MND in 2020 \textbf{10.1080/21678421.2019.1587629}. The total number of caregivers with available data at the 9-month timepoint (and for the change from baseline to 9 months was) was: ZBI, n = 66; EQ-5D, n = 67.

\subsection*{Caregiver and patient characteristics}
\autoref{tab_desc_stat} summarizes the baseline sociodemographic and clinical characteristics of the caregivers and patients. Caregivers had a mean (SD) age of 59.7 (11.8) years and most (70.2\%) caregivers were female, consistent with previous literature \parencite{schischlevskij_informal_2021}. Approximately half of the caregivers (4.8\%) reported being in paid employment.  Most caregivers (92.9\%) were direct family members of the patients and likely living together with a partner with MND as observed in previous studies \parencite{burke_longitudinal_2018, galvin_caregiving_2016, schischlevskij_informal_2021}. The median duration of caregiving (DOC) was three hours per day (as stated by the main CG; additional hours by more than one CG per patient are possible).  The mean ZBI scores at baseline was 21.1 (12.3). The EQ-VAS mean (79.8/100) was the same as the estimated EQ-5D-5L index score with 0.80/1. These HRQoL values were \textbf{lower/higher} than in the general British population (EQ-VAS mean = XX.X/100, SD = X.X [REF], EQ-5D-5L index value mean = X.X, SD = X.X) [REF].

@@

\subsection*{Impact of patient's functional status on caregivers}
The mean ZBI score by King's stage showed that caregivers patients with worse functional status had a greater burden, but results were not consistent. \autoref{fig:outcome-kings-stage} shows the ZBI score by King's stage at baseline, 6 months, and 9 months. The ZBI score was highest for carers of stage 4 patients, but higher for stage 2 compared to stage 3. Using baseline data only, scores were < 20 for stage 1/12 and >20 for stage 3/4. We found a weak, negative, correlation between the ZBI score and the ALSFRS-R ($\rho$ = −0.208, p < 0.002). Due to this inconsistency, functional status.

@@

\subsection*{Caregiver HRQoL and burden over time}
\autoref{tab_outcomes_time} The mean (SD) change in carers ZBI scores from baseline to 18 months was X.X (X.X) for the caregiver EQ-5D index score,X.X (X.X) for the caregiver EQ-5D VAS score, and X.X (X.X) for the ZBI score. For the two HRQoL measures, the change from baseline to 6 months or 9 months was not significant. We found significant differences in ZBI scores between 0 and 9  months (p value = X.XXX), mostly driven by the change ins cores beuween 0 and 6 months.

\autoref{fig:eq5d-domains-time} shows the EQ-5D domain scores over time. At baseline, no caregiver had any problem in self-care domain. %Examination of the EQ-5D domain scores (Fig. 1) showed that very few caregivers had extreme problems (<5\% for any domain) and the levels of problems were similar at baseline and at 18 months. At both time points, more caregivers had some problems in the domains of pain/discomfort (baseline: 42.8\%; 18 months: 46.8\%) and anxiety/depression (baseline: 30.3\%; 18 months: 33.9\%) than in the other three domains.

\begin{table}[H]
    \centering \singlespacing \small
    \caption{Means differences (95\% confidence intervals) and p-values from t-tests comparing of caregiver ZBI, EQ-VAS, and EQ-5D scores changes over time}
    \begin{tabular}{|L{6cm}|C{5cm}|C{2.5cm}|}
        \hline
        \PlainInput{tables/tab_outcomes_time}
    \end{tabular}
    \label{tab_outcomes_time}
    \caption*{\footnotesize 
                \textit{Notes:} CI, confidence interval; EQ-5D, EuroQol 5-dimension questionnaire; VAS, visual analog scale; ZBI, Zarit Burden Interview}
\end{table}

\begin{figure}[H]
    \centering
    \includegraphics[width=1\linewidth]{figures/eq5d-domains-time.png}
    \caption{Problems in the caregiver EQ-5D domains at baseline, 6 months, and 9 months.}
    \label{fig:eq5d-domains-time}
    \caption*{\footnotesize \textit{Notes:} MO, mobility; SC, self-care; UA, usual activities; PD, pain/depression; AD, anxiety/depression.}
\end{figure}

\subsection*{Relationship between ZBI and EQ-5D}
Correlations were weak or very weak between EQ-5D (index and VAS scores) and ZBI at each time point and for the change scores (\autoref{tab_outcomes_corr}). For the ZBI, the 18-month change score correlation was −0.16 (95\% confidence interval, CI: −0.22, −0.09) for the EQ-5D VAS score and −0.09 (95\% CI: −0.15, −0.03) for the EQ- 5D index score.

\begin{table}[H]
    \centering \singlespacing \small
    \caption{Spearman correlation coefficients between caregiver burden and quality of life scores at baseline, at 9 months, and change in scores between baseline and 9 months.}
    \begin{tabular}{|L{3.6cm}|C{1.7cm}|C{1.4cm}|C{1.7cm}|C{1.4cm}|C{1.7cm}|C{1.4cm}|}
        \hline
        \PlainInput{tables/tab_outcomes_corr}
    \end{tabular}
    \label{tab_outcomes_corr}
    \caption*{\footnotesize 
                \textit{Notes:} EQ-5D, EuroQol 5-dimension; EQ-VAS, EuroQol visual analog scale; ZBI, Zarit Burden Interview}
\end{table}

\subsection*{Factors influencing caregiver burden}
%As already shown above, the patient’s functional status had a clear influence on her/his caregiver’s burden. To provide a deeper analysis of further influencing factors on caregiver burden (ZBI score), multiple regression analysis was used (Table 3). The patient’s wheelchair dependency increased the ZBI score by 9.30 points. Additionally, a rise of 5.01 points was observed, if the patient needed permanent supervision. The main influence on the ZBI appeared to be the CG’s mental health impairment due to caregiving with an increase of 11.36 points. However, physical health impairment had no statistically significant impact on the ZBI score. Furthermore, patients’ age seemed to slightly lower the burden by −0.24 points for each increasing year. Neither the patient’s nor the CG’s gender had statistically significant impacts on the ZBI in the multivariate regression analysis.

\begin{table}[H]
    \centering \singlespacing \small
    \caption{Multivariable OLS regression models of associations between caregiver burden (ZBI) and caregiver- and patient-related independent variables}
    \begin{tabular}{|L{5cm}|R{1.6cm}|R{1.6cm}|R{1.6cm}|R{1.6cm}|R{1.6cm}|}
        \hline
        \PlainInput{tables/tab_multiOLS_zbi}
    \end{tabular}
    \label{tab_multiOLS_zbi}
    \caption*{\footnotesize 
                \textit{Notes:} ALSFRS-R, Revised Amyotrophic Lateral Sclerosis Functional Rating Scale; MQOL, McGill Quality of Life Questionnaire; ZBI, Zarit Burden Interview}
\end{table}

\begin{table}[H]
    \centering \singlespacing \small
    \caption{Multivariable OLS regression models of associations between health‐related quality of life (EQ-5D) and caregiver- and patient-related independent variables}
    \begin{tabular}{|L{5cm}|R{1.6cm}|R{1.6cm}|R{1.6cm}|R{1.6cm}|R{1.6cm}|}
        \hline
        \PlainInput{tables/tab_multiOLS_eq5d}
    \end{tabular}
    \label{tab_multiOLS_eq5d}
    \caption*{\footnotesize 
                \textit{Notes:} ALSFRS-R, Revised Amyotrophic Lateral Sclerosis Functional Rating Scale; EQ-5D, EuroQol 5-dimension; MQOL, McGill Quality of Life Questionnaire}
\end{table}

\clearpage
\section*{Supplementary tables}

\begin{table}[H]
    \centering \singlespacing \small
    \caption{Multivariable Beta regression models of associations between caregiver burden (ZBI) and caregiver- and patient-related independent variables}
    \begin{tabular}{|L{5cm}|R{1.6cm}|R{1.6cm}|R{1.6cm}|R{1.6cm}|R{1.6cm}|}
        \hline
        \PlainInput{tables/tab_multiBeta_zbi}
    \end{tabular}
    \label{tab_multiBeta_zbi}
    \caption*{\footnotesize 
                \textit{Notes:} ALSFRS-R, Revised Amyotrophic Lateral Sclerosis Functional Rating Scale; MQOL, McGill Quality of Life Questionnaire; ZBI, Zarit Burden Interview}
\end{table}

\begin{table}[H]
    \centering \singlespacing \small
    \caption{Univariable OLS regression models of associations, with ZBI as dependent variable, used in selection of independent variables for multivariable regression}
    \begin{tabular}{|L{6cm}|C{2cm}|C{3cm}|C{2cm}|}
        \hline
        \PlainInput{tables/tab_univarReg_zbi}
    \end{tabular}
    \label{tab_univarReg_zbi}
    \caption*{\footnotesize 
                \textit{Notes:} ALSFRS-R, Revised Amyotrophic Lateral Sclerosis Functional Rating Scale; HADS, Hospital Anxiety and Depression Scale; MQOL, McGill Quality of Life Questionnaire; ZBI, Zarit Burden Interview}
\end{table}

\begin{table}[H]
    \centering \singlespacing \small
    \caption{Univariable OLS regression models of associations, with EQ-5D as dependent variable, used in selection of independent variables for multivariable regression}
    \begin{tabular}{|L{6cm}|C{2cm}|C{3cm}|C{2cm}|}
        \hline
        \PlainInput{tables/tab_univarReg_eq5d}
    \end{tabular}
    \label{tab_univarReg_eq5d}
    \caption*{\footnotesize 
                \textit{Notes:} ALSFRS-R, Revised Amyotrophic Lateral Sclerosis Functional Rating Scale; HADS, Hospital Anxiety and Depression Scale; MQOL, McGill Quality of Life Questionnaire; ZBI, Zarit Burden Interview}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% %%%
%%% Bibliography %%%
